---
title: "aula2"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

#DataCamp

#INTRODUCTION TO R
# https://campus.datacamp.com/courses/free-introduction-to-r/chapter-1-intro-to-basics-1?ex=1  


#|##################|################################################################################
#                                             CAPÍTULO 1 - INTRO
#|##################|################################################################################


#Exercício 1
```{r}
# Calculate 3 + 4
3 + 4

# Calculate 6 + 12
6+12
```

#Exercício 2
```{r}
# An addition
5 + 5 

# A subtraction
5 - 5 

# A multiplication
3 * 5

 # A division
(5 + 5) / 2 

# Exponentiation
2^5

# Modulo
28%%6
```

#Exercício 3
```{r}
# Assign the value 42 to x
x <- 42

# Print out the value of the variable x
print (x)
```

#Exercício 4
```{r}
# Assign the value 5 to the variable my_apples
my_apples = 5

# Print out the value of the variable my_apples
print (my_apples)
```

#Exercício 5
```{r}
# Assign a value to the variables my_apples and my_oranges
my_apples <- 5
my_oranges <- 6

# Add these two variables together
my_apples + my_oranges

# Create the variable my_fruit
my_fruit = my_apples + my_oranges
```

#Exercício 6
```{r}
# Assign a value to the variable my_apples
my_apples <- 5 

# Fix the assignment of my_oranges
my_oranges <- 6 

# Create the variable my_fruit and print it out
my_fruit <- my_apples + my_oranges 
my_fruit
```

#Exercício 7
```{r}
# Change my_numeric to be 42
my_numeric <- 42

# Change my_character to be "universe"
my_character <- "universe"

# Change my_logical to be FALSE
my_logical <- FALSE
```

#Exercício 8
```{r}
# Declare variables of different types
my_numeric <- 42
my_character <- "universe"
my_logical <- FALSE 

# Check class of my_numeric
class(my_numeric)

# Check class of my_character
class(my_character)
```



#|##################|################################################################################
#                                            CHAPTER 2 - VECTORES
#|##################|################################################################################


#Exercício 1
```{r}
# Define the variable vegas
vegas <- "Go!"
```

#Exercício 2
```{r}
numeric_vector <- c(1, 10, 49)
character_vector <- c("a", "b", "c")

# Complete the code for boolean_vector
boolean_vector <- c(TRUE, FALSE, TRUE)
```

#Exercício 3
```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)
```

#Exercício 4
```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# Assign days as names of poker_vector
names(poker_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
```

#Exercício 5
```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# The variable days_vector
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
 
# Assign the names of the day to roulette_vector and poker_vector
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

names(poker_vector)
```

#Exercício 6
```{r}
A_vector <- c(1, 2, 3)
B_vector <- c(4, 5, 6)

# Take the sum of A_vector and B_vector
total_vector <- c(A_vector + B_vector )
  
# Print out total_vector
total_vector
```

#Exercício 7
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Assign to total_daily how much you won/lost on each day
total_daily <- c(poker_vector+roulette_vector)
```

#Exercício 8
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Total winnings with poker
total_poker <- sum(poker_vector)

# Total winnings with roulette
total_roulette <- sum(roulette_vector)

# Total winnings overall
total_week <- total_poker + total_roulette

# Print out total_week
  print(total_week)
```

#Exercício 9
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Calculate total gains for poker and roulette
total_poker <-sum(poker_vector)
total_roulette <- sum(roulette_vector)

# Check if you realized higher total gains in poker than in roulette 
print(total_poker)
print(total_roulette)
total_poker>total_roulette
```

#Exercício 10
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Define a new variable based on a selection
poker_wednesday <- poker_vector[3]
print(poker_wednesday)
```

#Exercício 11
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Define a new variable based on a selection
poker_midweek <- poker_vector[c(2,3,4)]
print(poker_midweek)
```

#Exercício 12
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Define a new variable based on a selection
roulette_selection_vector <- roulette_vector [c(2:5)]
```

#Exercício 13
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select poker results for Monday, Tuesday and Wednesday
poker_start <- poker_vector[c("Monday", "Tuesday", "Wednesday")]
  
# Calculate the average of the elements in poker_start
print(mean(poker_start))
```

#Exercício 14
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Which days did you make money on poker?
selection_vector <- poker_vector > 0
  
# Print out selection_vector
print(selection_vector)
print(selection_vector2 <- roulette_vector > 0)
```

#Exercício 15
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Which days did you make money on poker?
selection_vector <- poker_vector > 0

# Select from poker_vector these days
poker_winning_days <- poker_vector[c(selection_vector)]
print(poker_winning_days)
```

#Exercício 16
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Which days did you make money on roulette?
selection_vector <- roulette_vector> 0
print(selection_vector)

# Select from roulette_vector these days
roulette_winning_days <- roulette_vector[c(selection_vector)]
print(roulette_winning_days)
```



#|##################|################################################################################
#                                 CHAPTER 3 - MATRIZES
#|##################|################################################################################


#Exercício 1
```{r}
# Construct a matrix with 3 rows that contain the numbers 1 up to 9
matrix(c(1, 2, 3, 4, 5, 6, 7, 8, 9), byrow = TRUE, nrow=3)
#outra mas com a ordena??o por colunas
matrix(c(1:9), byrow = FALSE, ncol = 3)
```

#Exercício 2
```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Create box_office
box_office <- c(new_hope, empire_strikes, return_jedi)
print(box_office)

# Construct star_wars_matrix
star_wars_matrix <- matrix(box_office, byrow = TRUE, nrow = 3)
print(star_wars_matrix)

#outra matriz mas em coluna
#star_wars_matrix2 <- matrix(c(new_hope, empire_strikes, return_jedi), nrow = FALSE, ncol = 3)
```

#Exercício 3
```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Construct matrix
star_wars_matrix <- matrix(c(new_hope, empire_strikes, return_jedi), nrow = 3, byrow = TRUE)

# Vectors region and titles, used for naming
region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# Name the columns with region
colnames(star_wars_matrix) <- region

# Name the rows with titles
rownames(star_wars_matrix) <- titles
print(rownames)

# Print out star_wars_matrix
print(star_wars_matrix)
```

#Exercício 4
```{r}
# Construct star_wars_matrix
box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE,
                           dimnames = list(c("A New Hope", "The Empire Strikes Back", "Return of the Jedi"), 
                                           c("US", "non-US")))
print(star_wars_matrix)

# Calculate worldwide box office figures
worldwide_vector <- c(rowSums(star_wars_matrix))
print(worldwide_vector)

us_vector <- c(colSums(star_wars_matrix))
print(us_vector)
```

#Exercício 5
```{r}
# Construct star_wars_matrix
box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE,
                           dimnames = list(c("A New Hope", "The Empire Strikes Back", "Return of the Jedi"), 
                                           c("US", "non-US")))
print(star_wars_matrix)

# The worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)

# Bind the new variable worldwide_vector as a column to star_wars_matrix
all_wars_matrix <- cbind(star_wars_matrix,worldwide_vector)

print(worldwide_vector)
print(all_wars_matrix)
```

#Exercício 6

```{r}
# star_wars_matrix and star_wars_matrix2 are available in your workspace (ok mas aqui tenho que construir a 2 matriz)
star_wars_matrix

#números da box_office2
box_office2 <- c(475.5, 552.5, 310.7, 338.7, 380.3, 468.5)
star_wars_matrix2 <- matrix(box_office2, nrow = 3, byrow = TRUE, dimnames = (list(c("The Phantom Menace", "Attack of the Clones", "Revenge of Sith"), c("US", "Non-US"))))

#print inicial star_war_matrix2
print(star_wars_matrix2)

titles2 <- c("The Phantom Menace", "Attack of the Clones", "Revenge of Sith")

star_wars_matrix2 <- matrix(box_office2, nrow = 3, byrow = TRUE, dimnames = (list(titles2, c("US", "Non-US"))))
#print igual star_war_matrix2: é = construir com os nomes diretamente ou com a variável (vetor com os nomes)
print(star_wars_matrix2)

# Combine both Star Wars trilogies in one matrix
all_wars_matrix <- rbind(star_wars_matrix, star_wars_matrix2)

print(all_wars_matrix)
```

#Exercício 7
```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix
print(all_wars_matrix)

# Total revenue for US and non-US
total_revenue_vector <- colSums(all_wars_matrix)
  
# Print out total_revenue_vector
print(total_revenue_vector)
```

#Exercício 8
```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Select the non-US revenue for all movies
non_us_all <- all_wars_matrix[,2]
  
# Average non-US revenue
print(mean(non_us_all))
  
# Select the non-US revenue for first two movies
non_us_some <- all_wars_matrix[1:2,2]
  
# Average non-US revenue for first two movies
mean(non_us_some)
```

#Exercício 9
```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Estimate the visitors
visitors <- all_wars_matrix / 5
  
# Print the estimate to the console
print(visitors)
```

#Exercício 10
```{r}
# all_wars_matrix and ticket_prices_matrix are available in your workspace (ticket prices não estão)
all_wars_matrix

titles1_2 <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi", "The Phantom Menace", "Attack of the Clones", "Revenge of Sith")

ticket_prices <- c(5.0, 5.0, 6.0, 6.0, 7.0, 7.0, 4.0, 4, 4.5, 4.5, 4.9, 4.9)

ticket_prices_matrix <- matrix(ticket_prices, nrow = 6, byrow = TRUE, dimnames = list(titles1_2, region))

print(ticket_prices_matrix)


# Estimated number of visitors
visitors <- all_wars_matrix / ticket_prices_matrix

# US visitors
us_visitors <- visitors[,1]

# Average number of US visitors
mean(us_visitors)
```



#|##################|################################################################################
#                                         CHAPTER 4 - Fatores 
#|##################|################################################################################


#Exercício 1
```{r}
# Assign to the variable theory what this chapter is about!
theory <- "factors for categorical variables"
```

#Exercício 2
```{r}
# Sex vector
sex_vector <- c("Male", "Female", "Female", "Male", "Male")

# Convert sex_vector to a factor
factor_sex_vector <- factor(sex_vector)

# Print out factor_sex_vector
print(factor_sex_vector)
```

#Exercício 3
```{r}
# Animals
animals_vector <- c("Elephant", "Giraffe", "Donkey", "Horse")
factor_animals_vector <- factor(animals_vector)
factor_animals_vector

# Temperature
temperature_vector <- c("High", "Low", "High","Low", "Medium")
factor_temperature_vector <- factor(temperature_vector, order = TRUE, levels = c("Low", "Medium", "High"))
factor_temperature_vector
```

#Exercício 4
```{r}
# Code to build factor_survey_vector
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
print(factor_survey_vector)

# Specify the levels of factor_survey_vector
levels(factor_survey_vector) <- c("Female", "Male")

factor_survey_vector
```

#Exercício 5
```{r}
# Build factor_survey_vector with clean levels
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
levels(factor_survey_vector) <- c("Female", "Male")
factor_survey_vector

# Generate summary for survey_vector
summary(survey_vector)

# Generate summary for factor_survey_vector
summary(factor_survey_vector)
```

#Exercício 6
```{r}
# Build factor_survey_vector with clean levels
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
levels(factor_survey_vector) <- c("Female", "Male")

# Male
male <- factor_survey_vector[1]

# Female
female <- factor_survey_vector[2]

# Battle of the sexes: Male 'larger' than female?
male > female
```

#Exercício 7
```{r}
# Create speed_vector
speed_vector <- c("medium", "slow", "slow", "medium", "fast")
print(speed_vector)
```

#Exercício 8
```{r}
# Create speed_vector
speed_vector <- c("medium", "slow", "slow", "medium", "fast")

# Convert speed_vector to ordered factor vector
factor_speed_vector <- factor(speed_vector, ordered = TRUE, levels = c("slow", "medium", "fast"))

# Print factor_speed_vector
print(factor_speed_vector)
summary(factor_speed_vector)
```

#Exercício 9
```{r}
# Create factor_speed_vector
speed_vector <- c("medium", "slow", "slow", "medium", "fast")
factor_speed_vector <- factor(speed_vector, ordered = TRUE, levels = c("slow", "medium", "fast"))

# Factor value for second data analyst
da2 <- factor_speed_vector[2]
print(da2)

# Factor value for fifth data analyst
da5 <- factor_speed_vector[5]

# Is data analyst 2 faster than data analyst 5?
da2 > da5
```


#|##################|################################################################################
#                               CHAPTER 5 - DATA FRAMES (quadros de dados)
#|##################|################################################################################


#Exercício 1
```{r}
# Print out built-in R data frame
mtcars
```

#Exercício 2
```{r}
# Call head() on mtcars
head(mtcars)
# Call tail() on mtcars - este foi para ver os ?ltimos
tail(mtcars)
```

#Exercício 3
```{r}
# Investigate the structure of mtcars
str(mtcars)
```

#Exercício 4
```{r}
#Creating a data frame
name <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
type <- c("Terrestrial Planet", "Terrestrial Planet", "Terrestrial Planet", "Terrestrial Planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)

# Create a data frame from the vectors
planets_df <- data.frame(name, type, diameter, rotation, rings)
print(planets_df)
```

#Exercício 5
```{r}
# Check the structure of planets_df
str(planets_df)
```

#Exercício 6
```{r}
# The planets_df data frame from the previous exercise is pre-loaded

# Print out diameter of Mercury (row 1, column 3)
print(planets_df[1,3])

# Print out data for Mars (entire fourth row)
print(planets_df[4,])

```

#Exercício 7
```{r}
# Select first 5 values of diameter column
print(planets_df)
print(planets_df[1:5,3])
```

#Exercício 8
```{r}
# Select the rings variable from planets_df
rings_vector <- planets_df[,5]
  
# Print out rings_vector
print(rings_vector)
```

#Exercício 9
```{r}
# Adapt the code to select all columns for planets with rings
planets_df[rings_vector,]
```

#Exercício 10
```{r}
# Select planets with diameter < 1
subset(planets_df, subset = diameter<1)
# Select planets with rings
subset(planets_df, subset = rings)
```

#Exercício 11
```{r}
# Play around with the order function in the console
a <- c(120, 80, 500)
print(a)
order(a)
a[order(a)]

```

#Exercício 12
```{r}
# Use order() to create positions
positions <- order(planets_df$diameter)
print(positions)

# Use positions to sort planets_df
planets_df[positions, ]

```



#|##################|################################################################################
#                               CHAPTER 6 - LISTS
#|##################|################################################################################


#1
#Congratulations! At this point in the course you are already familiar with:

Vectors (one dimensional array): can hold numeric, character or logical values. The elements in a vector all have the same data type.

Matrices (two dimensional array): can hold numeric, character or logical values. The elements in a matrix all have the same data type.

Data frames (two-dimensional objects): can hold numeric, character or logical values. Within a column all elements have the same data type, but different columns can be of different data type.

#2
#A list in R is similar to your to-do list at work or school: the different items on that list most likely differ in length, characteristic, and type of activity that has to be done.

A list in R allows you to gather a variety of objects under one name (that is, the name of the list) in an ordered way. These objects can be matrices, vectors, data frames, even other lists, etc. It is not even required that these objects are related to each other in any way.

You could say that a list is some kind super data type: you can store practically any piece of information in it!


#Exercício 3
```{r}
# Vector with numerics from 1 up to 10
my_vector <- 1:10 
my_vector

# Matrix with numerics from 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)
my_matrix

# First 10 elements of the built-in data frame mtcars
my_df <- mtcars[1:10,]
my_df

# Construct list with these different elements:
my_list <- list(my_vector, my_matrix, my_df)
my_list

```

#Exercício 4
```{r}
# Vector with numerics from 1 up to 10
my_vector <- 1:10 

# Matrix with numerics from 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)

# First 10 elements of the built-in data frame mtcars
my_df <- mtcars[1:10,]

# Adapt list() call to give the components names
my_list <- list(my_vector, my_matrix, my_df)
names(my_list) <- c("vec", "mat", "df")

# Print out my_list
print(my_list)
```

#Exercício 5
```{r}
# The variables mov, act and rev are available - aqui tive que as criar de acordo com o que estava

mov = "The Shining"

act = c("Jack Nicholson", "Shelley Duvall", "Danny Lloyd", "Scatman Crothers", "Barry Nelson")

rev = data.frame(c(4.5, 4.0, 5.0), c("IMDb1", "IMDb2", "Imdb3"), c("Best Horror Film I Have Ever Seen", "Best Horror Film I Have Ever Seen", "A masterpiece of psychological horror"))
row.names(rev) = c(1,2,3)
colnames(rev) = c("scores", "sources", "comments")

mov
act
rev

# Finish the code to build shining_list
shining_list <- list(mov, act, rev)

names(shining_list) <- c("moviename","actors" ,"reviews")

print(shining_list)


```

#Exercício 6
```{r}
# Print out the vector representing the actors
shining_list$actors

# Print the second element of the vector representing the actors
shining_list$actors[2]

```

#Exercício 7
```{r}
# shining_list, the list containing movie name, actors and reviews, is pre-loaded in the workspace

# We forgot something; add the year to shining_list
shining_list_full <- c(shining_list, year = 1980)
shining_list_full

# Have a look at shining_list_full
str(shining_list_full)


##########################################################################

# INCLUINDO TAMBÉM POR EXEMPLO A VARIÁVEL month
shining_list_full <- c(shining_list, year = 1980, month = c(15,25,35))
shining_list_full
shining_list_full$month1

# Have a look at shining_list_full
str(shining_list_full)


```








